Data-Driven Multi-grid Solver for Accelerated \\ Pressure Projection

### 2021-10-19 tex/article.tex
Updates from Overleaf

\begin{figure}
    \caption{Sketch of a Convolutional Neural Networks (CNN) and/or a Multi-grid V-cycle.}
    \label{fig:multigrid}
\end{figure}

@@

The linearity of equation \ref{eq:aer} is a critical property which the data-driven method must maintain. Not only will this increase generality, but a nonlinear method could not be used iteratively since the modes of the residual will change as the solver converges. However, this linearity is only with respect to the update and residual fields. The operators of the MG method can all potentially be made into parameterized nonlinear functions of $A$, embedding information about the discrete problem and boundary conditions.

Which MG operation is the best candidate for such an approach? The prolongation operator (and its transpose restriction operator) is not inherently a function of $A$, and while related super-resolution operators have seen use in the literature, these are nonlinear functions of the field. The precondition operator is an option, but a simple Jacobi preconditioner
\begin{equation}
    \epsilon = D^{-1}r
\end{equation}
where $D$ is the diagonal of $A$, is as fast as possible and is sufficient for the V-cycle to distribute the residual throughout the domain. Therefore, for the remainder of the paper we will focus on the smoother as the best candidate for parameterization, using a simple Jacobi preconditioner and uniform pooling/distribution for restriction/prolongation.

@@

There are few constraints on $\tilde A^{-1}$: it must have units of $A^{-1}$, and symmetry of $A$ implies the approximate inverse should also be symmetric. For simplicity, the diagonal and off-diagonal coefficients of $\tilde A^{-1}$ are constructed independently 
\begin{equation}\label{eq:approxinv}
    \tilde a^{-1}_{ii} = \frac{f_d(a_{ii}/s\,|\theta_d)}{a_{ii}} , \quad
    \tilde a^{-1}_{ij} = \frac{f_o(a_{ij}/s\,|\theta_o)}{a_{ii}+a_{jj}}
\end{equation}
where function inputs are scaled by the maximum off-diagonal $s=\max(A-D)$ and the outputs are scaled by the diagonal elements. Note a Jacobi smoother would use a diagonal function $f_d=1$ and off-diagonal function $f_o=0$. With the learning problem now reduced to two normalized single-variable functions, the specific choice of parameterization is not critical and a simple quadratic polynomial $f(a) = \sum_{p=0}^2 a^p \theta_p$ is chosen for $f_o$ and $f_d$. Using higher-order polynomials, splines, and interpolating kernels did not significantly change the results.

@@

Finally, the acceleration of the `union' data-driven smoother is evaluated on new example data of a different size, $n=64^M$ points. Figure~\ref{fig:synthetic time} shows the time to reduce the residual of each case by $10^{-3}$ relative to the time to run a single V-cycle using a Jacobi smoother. While a Jacobi smoother requires  hundreds or even thousands of V-cycles to converge, the Jacobi-like parameterized smoother is only around 30-50\% slower per V-cycle, and yet converges in only 1-3 V-cycles in all cases. This results in a 80-174\% speed-up (mean: 119\%) relative to optimized serial Gauss-Sidel and SOR smoothers. This speed-up will be even more favorable in parallel computation as the new smoother operates directly on the local residual without back-propagation, eliminating any additional communication.

\begin{figure}
    \centering
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \caption{circle flow}
        \label{fig:circle}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \caption{Taylor-Green Vortex}
        \label{fig:TGV}
    \end{subfigure}
    \begin{subfigure}[b]{0.3\textwidth}
        \centering
        \caption{flapping wing}
        \label{fig:wing}
    \end{subfigure}
    % \begin{minipage}{0.47\textwidth}
        \begin{subfigure}{0.47\textwidth}
        \centering
        \caption{torus flow}
        \label{fig:donut}
        \end{subfigure}
        \hfill
        \begin{subfigure}{0.47\textwidth}
        \centering
        \caption{swimming shark}
        \label{fig:shark}
        \end{subfigure}
    % \end{minipage}
    \hfill
    % \begin{subfigure}[b]{0.47\textwidth}
    %     \centering
    %     \includegraphics[width=\textwidth]{figures/compareloss.png}
    %     \caption{single-cycle residual reduction}
    %     \label{fig:simulation residual}
    % \end{subfigure}
    \caption{(a-e) Snapshots from the five flow simulation cases. The Taylor-Green Vortex and torus flow are 3D, the flapping wing is dynamic without a background flow and the swimming shark is a deforming geometry.}
    \label{fig:simulation cases}
\end{figure}

@@

Finally, the parameterized $f_d,\,f_o$ functions used in the approximate inverse equation \ref{eq:approxinv} are shown in Figure \ref{fig:tuned inverse}. It is interesting to note that the diagonal functions are all somewhat centered on one and the off-diagonal values all have an essentially zero intercept as these are the values for the Jacobi smoother. We also note that the `wing' and `shark' cases produce parameterizations which are similar to each other but significantly different than the others cases. This is reasonable as dynamic and deforming geometries put unique burdens on the pressure projection step, as shown in the longer convergence times for these cases shown in Figure \ref{fig:simulation time}. Adopting a data-driven and auto-tuned approach enables these pressure-projection dominated cases to achieve significant accelerations.

@@

The potential of machine learning and computer science advances to improve fluid dynamics is vast, but well-applied classical methods and constraints only helps to focus this effort. Wherever possible, this work has made the simplest choice in parameterization, leaving significant opportunities for future improvements. Many other flow solvers and linear systems are yet to be explored, and extensions to this basic data-driven approach could accelerate solutions for a wide array of discrete partial differential equations.

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{figures/drawing}
    \caption{Sketch of a Convolutional Neural Networks (CNN) encoder/decoder and/or a Multi-grid V-cycle. Data is restricted down to a reduced size for faster solving before being pushed back up to the full size. A single reduction is shown, but this process repeats recursively.}
    \label{fig:multigrid}
\end{figure}

@@

The linearity of equation \ref{eq:aer} is a critical property which the data-driven method must maintain. Consider that the modes of the residual will change as the solver converges for even a single project case, meaning that a nonlinear method could not be applied iteratively. However, this linearity is only with respect to the update and residual fields. The operators of the MG method can all potentially be made into parameterized nonlinear functions of $A$, embedding information about the discrete problem and boundary conditions.

Which MG operation is the best candidate for such an approach? There is certainly scope for optimization in the prolongation operator (and its transpose restriction operator) in algebraic multi-grid methods.

is not inherently a function of $A$, and while related super-resolution operators have seen use in the literature, these are nonlinear functions of the field. The precondition operator is an option, but a simple Jacobi preconditioner
\begin{equation}
    \epsilon = D^{-1}r
\end{equation}
where $D$ is the diagonal of $A$, is as fast as possible and is sufficient for the V-cycle to distribute the residual throughout the domain. Therefore, for the remainder of the paper we will focus on the smoother as the best candidate for parameterization, using a simple Jacobi preconditioner and uniform pooling/distribution for restriction/prolongation.

@@

There are few constraints on $\tilde A^{-1}$: it must have units of $A^{-1}$, and symmetry of $A$ implies the approximate inverse should also be symmetric. For simplicity, the diagonal and off-diagonal coefficients of $\tilde A^{-1}$ are constructed independently 
\begin{equation}\label{eq:approxinv}
    \tilde a^{-1}_{ii} = \frac{f_d(a_{ii}/s\,|\theta_d)}{a_{ii}} , \quad
    \tilde a^{-1}_{ij} = \frac{f_o(a_{ij}/s\,|\theta_o)}{a_{ii}+a_{jj}}
\end{equation}
where function inputs are scaled by the maximum off-diagonal $s=\max(A-D)$ and the outputs are scaled by the diagonal elements. Note a Jacobi smoother would use a diagonal function $f_d=1$ and off-diagonal function $f_o=0$ and that $f_0(0)=0$ is required to maintain the sparsity of $\tilde A^{-1}$. With the learning problem now reduced to two normalized single-variable functions, the specific choice of parameterization is not critical and a simple cubic is chosen for $f_d$ and $f_o$. Using higher-order polynomials, splines, and interpolating kernels did not significantly change the results.

@@

Finally, the acceleration of the `union' data-driven smoother is evaluated on new example data of a different size, $n=64^M$ points. Figure~\ref{fig:synthetic time} shows the time to reduce the residual of each case by $10^{-3}$ relative to the time to run a single V-cycle using a Jacobi smoother. The Jacobi smoother is not sufficient to solve even these simple synthetic problems rapidly, requiring many hundreds o V-cycles. Yet, the Jacobi-like parameterized smoother is only around 30-50\% slower per V-cycle, and converges in only 1-3 V-cycles in all cases. This results in a 80-210\% speed-up (mean: 133\%) relative to optimized serial Gauss-Sidel and SOR smoothers. This speed-up will be even more favorable in parallel computation as the new smoother operates directly on the local residual without back-propagation, eliminating any additional communication.

\begin{figure}
    \centering
    \begin{subfigure}[a]{\textwidth}
        \centering 
        \includegraphics[width=\textwidth]{figures/circletriple.png}
        \caption[-1mm]{Circle flow}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/donuttriple.png}
        \caption{Torus flow (slice)}
    \end{subfigure}
    \begin{subfigure}[c]{0.6\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/TGVtriple.png}
        \caption{Taylor-Green Vortex (slice)}
    \end{subfigure}
    \begin{subfigure}[c]{0.6\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/wingtriple.png}
        \caption{Flapping wing}
    \end{subfigure}
    \begin{subfigure}[b]{\textwidth}
        \centering
        \includegraphics[width=\textwidth]{figures/sharktriple.png}
        \caption{Swimming shark}
    \end{subfigure}
    \hfill
    \caption{Snapshots from the five flow simulation cases showing the pressure solution (left), the initial residual (middle), and the residual after a single V-cycle (right). The Taylor-Green Vortex and torus flow are 3D, the flapping wing is dynamic without a background flow and the swimming shark is a deforming geometry. All residuals are visualized using $\log_{10}\text{abs}(r)$ on the same scale, $10^{-6} \ldots 10^{-1}$.}
    \label{fig:simulation cases}
\end{figure}

@@

Finally, the parameterized $f_d,\,f_o$ functions used in the approximate inverse equation \ref{eq:approxinv} are shown in Figure \ref{fig:tuned inverse}. It is interesting to note that the diagonal functions are all somewhat centered on one, the value for the Jacobi smoother. We also note that the `wing' and `shark' cases produce parameterizations which are similar to each other but significantly different than the others cases. This is reasonable as dynamic and deforming geometries put unique burdens on the pressure projection step, as shown in the longer convergence times for these cases shown in Figure \ref{fig:simulation time}. Adopting a data-driven and auto-tuned approach enables these pressure-projection dominated cases to achieve significant accelerations.

@@

The potential of machine learning advances to improve fluid dynamics is vast, but well-applied classical methods and constraints are needed to focus this potential. Wherever possible, this work has made the simplest choice in parameterization, leaving significant opportunities for future improvements. Many other flow solvers and linear systems are yet to be explored, and extensions to this basic data-driven approach could accelerate solutions for a wide array of discrete partial differential equations.
